---
title: "TG_SCDi_changes"
author: "Dan Ackerman"
date: "December 10, 2015"
output: html_document
---

Trying to look at changes in saturation of TG species to ultimately compare these changes to those seen in PC and other PL species

```{r}
## creating dataframe

setwd("C:/R/Aug2015")
dat1<-read.table("Input_data_InVitro_0.5%_DMSO_DOXvsVehicle.csv", header=TRUE, sep=",")
dat2<-read.table("Input_data_InVitro_0.5%_SCDi_DOXvsVehicle.csv", header=TRUE, sep=",")

rep1<-read.table("Input_data_InVitro_5%_DMSO_DOXvsVehicle.csv", header=TRUE, sep=",")
rep2<-read.table("Input_data_InVitro_5%_SCDi_DOXvsVehicle.csv", header=TRUE, sep=",")

dat3<-merge(dat1, dat2, by="Label")
rep3<-merge(rep1, rep2, by="Label")
colnames(dat3)<-c(colnames(dat3)[1], paste0("S-", colnames(dat3[2:21])))
colnames(rep3)<-c(colnames(rep3)[1], paste0("R-", colnames(rep3[2:20])))


comb1<-merge(dat3, rep3, by="Label")



library(stringr)


######################################################
## Calculating TG sat distribution
tgs1<-comb1[grep("TG", comb1[,1]),]


tg_labels<-as.data.frame(str_match(tgs1[,1], "^(.*):(.*)/(.*):(.*)/(.*):(.*)"))
## http://stackoverflow.com/questions/24100613/split-column-into-multiple-columns-r
colnames(tg_labels)<-c("species", "ch1", "ch1DBs", "ch2", "ch2DBs", "ch3", "ch3DBs")
##adds column names
##cleaning up
tg_labels$ch1<-gsub("TG\\(", "", tg_labels$ch1)
tg_labels$ch3DBs<-gsub("\\)", "", tg_labels$ch3DBs)

tgs2<-merge(tgs1,tg_labels, by.x="Label", by.y = "species")

tgs2$ch1DBs<-as.numeric(as.character(tgs2$ch1DBs))
tgs2$ch2DBs<-as.numeric(as.character(tgs2$ch2DBs))
tgs2$ch3DBs<-as.numeric(as.character(tgs2$ch3DBs))


tgs2$unsat<-tgs2$ch1DBs+tgs2$ch2DBs+tgs2$ch3DBs
## reorder

tgs2<- tgs2[,c(1, 41:47, 2:40)]

require(data.table)
dat4<-as.data.frame(setDT(tgs2)[,lapply(.SD, sum), by=unsat, .SDcols=-c('Label',"ch1", "ch1DBs", "ch2", "ch2DBs", "ch3", "ch3DBs")])

```

Function to divide each column by column total. 

```{r, warning=FALSE, error=FALSE, echo=FALSE}
scal<- function(dat){
  DAT<-dat
  nms<- colnames(DAT)
  tgs_col_sum<-apply(DAT[,-1],2,sum)  #calculating column total
  df<-data.frame(DAT[,1])
  df2<-sweep(DAT[,-1],2,tgs_col_sum, "/")   #each dataframe element divided by column total

  df3<-cbind(df,df2)
  colnames(df3)<- nms
  df3[order(df3[,1]),]
}

```

Scaling:

```{r}
dat5<- scal(dat4)

nmes<- gsub("_", "", colnames(dat5)) 
colnames(dat5)<- gsub(" ", "", nmes)  
```

Creating a dataframes that contain effect of SCDi under various conditions 
Split data into pairs of veh vs SCDi under various conditions then subtract mean of Veh from all values. 
```{r}

##mnsub() outputs function that calculates mean of control group and subtracts it from all values
mnsub<- function(df1, columns){
  
  df2  <-df1[,grep(columns, colnames(df1))]
  
  function(grepstring){
    x<- apply(df2[,grep(grepstring, colnames(df2))],1,mean)
    df3<- sweep(df2,1,x, "-")
    df3<- cbind(data.frame(DBs=dat4[,1]), df3)
    require(tidyr)
    long_df<- gather(df3, Sample, Measurement, 2:length(colnames(df3)))
    long_df$Group<-unlist(strsplit(as.character(long_df$Sample), '[1-5]$'))
    long_df
  }
}

.s_std<- mnsub(dat5, "S-STD")
  S_STD<- .s_std("Veh")
.s_dox<- mnsub(dat5, "S-DOX")
  S_DOX<- .s_dox("Veh")
.r_std<- mnsub(dat5, "R-STD")
  R_STD<- .r_std("Veh")
.r_dox<- mnsub(dat5, "R-DOX")
  R_DOX<- .r_dox("Veh")
```

Now plotting boxplot of mean subtracted values. All "Veh" samples should be clustering around 0.

```{r}
require(ggplot2)
plotobj<- function(df1){
  ggplot(df1, aes(x=factor(DBs), y=Measurement, fill= factor(Group)))+    geom_boxplot()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  labs(x="Double Bonds") +
    coord_cartesian(ylim = c(-0.2, 0.3))
}

a<- plotobj(S_STD)
  a+ggtitle("TG species, 0.5% serum, no DOX, effect of SCDi")
b<- plotobj(S_DOX)
  b+ggtitle("TG species, 0.5% serum, with DOX, effect of SCDi")
c<- plotobj(R_STD)
  c+ggtitle("TG species, 5% serum, no DOX, effect of SCDi")
d<- plotobj(R_DOX)
  d+ggtitle("TG species, 5% serum, with DOX, effect of SCDi")
```


Identifying outliers

```{r, error=FALSE, warning=FALSE}
id_outl<- function(df){
  ggplot(df, aes(x=factor(DBs), y=Measurement, color=Sample))+
  geom_point(position = position_jitter(w = 0.2), stat = "identity")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +   
  labs(x="Double Bonds")+ facet_grid(.~Group)
}

e<- id_outl(S_STD)
  e+ ggtitle("TG species, 0.5% serum, no DOX, effect of SCDi")
f<- id_outl(S_DOX)
  f+ggtitle("TG species, 0.5% serum, with DOX, effect of SCDi")
g<- id_outl(R_STD)
  g+ggtitle("TG species, 5% serum, no DOX, effect of SCDi")
h<- id_outl(R_DOX)
  h+ggtitle("TG species, 5% serum, with DOX, effect of SCDi")
```


```{r, error=FALSE, warning=FALSE}  
  ##mneffect() outputs function that calculates mean of control group and subtracts it from treatment values. 
mneffect<- function(df1, columns){
  #"columns" argument is character string that is present in column names of columns that should be included
  #"columns" string will be searched for using 'grep()' function
  #output is another function which will utilize dataframe to generate ggplot
  
  df2  <-df1[,grep(columns, colnames(df1))]
  
  function(grepstring="Veh"){
    x<- apply(df2[,grep(grepstring, colnames(df2))],1,mean)
    df3<- sweep(df2[,-grep(grepstring, colnames(df2))],1,x, "-")
    
    df4<- cbind(data.frame(DBs=df1[,1]), df3)
      df4$mean<- apply(df3[,2:length(colnames(df3))],1,mean)
        df4$sd<- apply(df3[,2:length(colnames(df3))],1,sd)
    df4$ctrlmn<- x
    z<- ggplot(df4, aes(x=DBs, y=mean, fill=mean>0))+
      geom_bar(stat="identity", position="identity", color="black", size=0.5)+
      geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9))+
      scale_x_continuous(breaks=seq(0,7,1))+
      scale_y_continuous(breaks=seq(-0.1,0.3,0.05))+
      scale_fill_manual(values=c("#CCEEFF", "#FFDDDD"), guide=FALSE) +
      geom_hline(aes(yintercept=0))+labs(y="relative change")+
      coord_cartesian(xlim = c(-0.5, 7.5), ylim=c(-.2, 0.45))+
      theme(axis.title.x=element_text(size = 25), axis.text.x=element_text(size=30),
        axis.title.y=element_text(size = 25), axis.text.y=element_text(size=30))+
      theme_bw()+
  
    geom_point(aes(size=ctrlmn, x= DBs, y=0.35), 
               shape = 21, color= "black", fill= "cornsilk")+
        scale_size_area(max_size=6.5, guide=FALSE)+
        annotate("text", x=3.5, y=0.43, label= "TG species", size=5)+
        annotate("text", x=3.5, y=0.39, label= "Abundance", size=3.5)
  z  
    
  }
}

  


.s_std_chng<- mneffect(dat5, "S-STD")
  S_STD_chng<- .s_std_chng("Veh")
.s_dox_chng<- mneffect(dat5, "S-DOX")
  S_DOX_chng<- .s_dox_chng("Veh")
.r_std_chng<- mneffect(dat5, "R-STD")
  R_STD_chng<- .r_std_chng("Veh")
.r_dox_chng<- mneffect(dat5, "R-DOX")
  R_DOX_chng<- .r_dox_chng("Veh")
  
  
S_STD_chng+ annotate("text", x=3.5, y=0.3, 
                 label= "TG species, 0.5% serum, no DOX, effect of SCDi", size=3.5)
S_DOX_chng+ annotate("text", x=3.5, y=0.3, 
                 label= "TG species, 0.5% serum, with DOX, effect of SCDi", size=3.5)
R_STD_chng+ annotate("text", x=3.5, y=0.3, 
                 label= "TG species, 5% serum, no DOX, effect of SCDi", size=3.5)
R_DOX_chng+ annotate("text", x=3.5, y=0.3, 
                 label= "TG species, 5% serum, with DOX, effect of SCDi", size=3.5)


```
  